---
title: "Axis_analysis_18_feb_2019"
author: "Kevin Healy"
date: "18 feb 2019"
output:
  pdf_document: default
  html_document: default
---

This script uses the metrics estimated in the previous scripts to run a series of `MCMCglmm` models in order to fit phylogentic regressions for each of the life history metrics against body mass and the matrix dimension. As each regression is run over 100 phylogentic trees using the mulTree package the output of each model is exported out of the R enviornment into the current directory. The required elements can then be read in as needed for the rest of the analysis. For more on the `multree` package see https://github.com/TGuillerme/mulTree

##Note running this script in full will result in a runtime of approximatly 1 week.


First we load the required packages including MCMCglmm and the mulTree package which runs the MCMCglmm models over the 100 phylogenies.
```{r source functions, message = FALSE, results='hide', warning = FALSE}

library(phytools)
library(MCMCglmm)
library(mulTree)

source("Demography_functions.R")
```

Next we upload the datafile which has the life history metrics calculated from the previous Pop_metric_calulation script.

```{r  data}

pop_data <- read.csv("axis_analysis_data.csv", 
                         sep = ",", header = T)

```

Next we upload the matching distribution of phylogenies calculated from the previous Phylogeny_construction and Pop_metric_calulation scripts. We do a quick ultrametric check and then plot out our tree.

```{r phylogeny, message = FALSE, warning = FALSE, fig.align ='center'}

axis_trees <- read.tree("axis_analysis_phylo.tre")


#check if ultrametric
is_ultra <- vector()
for(i in 1:100) { 
  is_ultra[i] <- is.ultrametric(axis_trees[[i]])
  }
ultra_row <- which(is_ultra ==TRUE)

axis_trees <- axis_trees[ultra_row]

plot(axis_trees[[1]], cex = 0.2, type = "fan")


```



We then apply a log10 transform to each of the life histroy metrics.


```{r log data,  message = FALSE, warning = FALSE}

log_list <- c("life_time_La",
              "mean_repo_rate_stable_state",
              "mean_repo_rate",
              "gen_time",
              "M_rep_lif_exp",
              "mass_g",
              "matrix_size")

pop_data_log <- pop_data

pop_data_log[,log_list] <- sapply(pop_data[,log_list], function(x) log10(x))

```

And mean center the data

```{r mean center data }
mean_c_list  <- c( "life_time_La", 
                  "mean_repo_rate",
                  "gen_time",
                  "M_rep_lif_exp",
                  "matrix_size",
                  "gini",
                  "mean_repo_rate_stable_state",
                  "mxlxsd",
                  "surv_sd",
                  "mass_g")

pop_data_log_mc <- pop_data_log
pop_data_log_mc[,mean_c_list] <- sapply(pop_data_log[,mean_c_list], function(x) mean_center(x))

```


#MCMCglmm

First we make a multree object that holds both the data and the multiphylo object. We also specify the random terms which in this case are the animal term for the phylogentic random effect and species for withn speices population level variation.
```{r mulTree}
pop_multree <- as.mulTree(data = pop_data_log_mc, tree = axis_trees, taxa = "animal", rand.terms = ~animal + species)

```

Now now also set a prior
```{r MCMC prior}

prior<-list(R = list(V = 1/2, nu=0.002), 
            G = list(G1=list(V = 1/2,n = 1, alpha.mu=rep(0,1), alpha.V= diag(1)*10^3), 
                     G1=list(V = 1/2,n = 1, alpha.mu=rep(0,1), alpha.V= diag(1)*10^3)))

```


We then set the number of iterations, thinning and the burnin to 1100000, 500 and 100000 as these have previously been found to give outputs where the chains converge and where the effective sample size is more then 1000. These criterion are automatcally checked as MulTree runs each model and are stored in the files ending _conv.rda

For the of running this script we run a dummy veresion with much lower parameters. For the full run use thr parameters above.

```{r MCMC parameters}

#parameters <- c(1100000, 500, 100000)
parameters <- c(2200, 1, 200)


```


Now we run each sets of models. The effective sample size (ESS) is set to 1000 and the number of chains to 2. This applies to each phylogeny so there will be 200 chains for each metric per run. 

#Age at first reproduction

```{r La allometries mulTree run}
formula_la <- life_time_La ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_la,
        priors = prior,
        parameters = parameters,
        output = "la_run",
        ESS = 1000,
        chains = 2)
```


#Mean reproductive rate with the population at a stable state distribution.


```{r mean_repo_rate allometries mulTree run}
formula_mean_repo_rate <- mean_repo_rate_stable_state ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_mean_repo_rate,
        priors = prior,
        parameters = parameters,
        output = "mean_repo_rate_run",
        ESS = 1000,
        chains = 2)
```



#Mean reproductive rate with the population not at a stable state distribution.


```{r mean_repo_rate_st allometries mulTree run}
formula_mean_repo_rate_nst <- mean_repo_rate ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_mean_repo_rate_nst,
        priors = prior,
        parameters = parameters,
        output = "mean_repo_rate_nst_run",
        ESS = 1000,
        chains = 2)
```



#Generation Time

```{r gen allometries mulTree run}
formula_gen_time <- gen_time ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_gen_time,
        priors = prior,
        parameters = parameters,
        output = "gen_time_run",
        ESS = 1000,
        chains = 2)
```


#Life expectancy conditional on reaching sexual maturity


```{r M_rep_lif_exp allometries mulTree run}
formula_M_rep_lif_exp <- M_rep_lif_exp ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_M_rep_lif_exp,
        priors = prior,
        parameters = parameters,
        output = "M_rep_lif_exp_run",
        ESS = 1000,
        chains = 2)
```


#Reporductive spread as measured using the gini index


```{r gini allometries mulTree run}
formula_gini <- gini ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_gini,
        priors = prior,
        parameters = parameters,
        output = "gini_run",
        ESS = 1000,
        chains = 2)
```


#Standard deviation of reporduction as measured using mxlx


```{r mxlxsd allometries mulTree run}
formula_mxlxsd <- mxlxsd ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_mxlxsd,
        priors = prior,
        parameters = parameters,
        output = "mxlxsd_logged_run",
        ESS = 1000,
        chains = 2)
```


#Standard deviation of mortality rates

```{r surv_sd allometries mulTree run}
formula_surv_sd <- surv_sd ~ mass_g + matrix_size

mulTree(mulTree.data = pop_multree,
        formula = formula_surv_sd,
        priors = prior,
        parameters = parameters,
        output = "surv_sd_run",
        ESS = 1000,
        chains = 2)
```


